Individual heterogeneity in the functional network topography of medial prefrontal cortex

### 2019-05-23 Manuscript.Rmd
halfway through the discussion

---
title: 'Individual heterogeneity in the functional network topography of medial prefrontal cortex'
author: "Claudio Toro-Serey^1^, Sean M. Tobyne^2^, & Joseph T. McGuire^1^"
output:
  pdf_document: 
    keep_tex: false
    highlight: haddock
bibliography: mPFC_paper.bib
csl: apa.csl
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
    - \linenumbers
    - \usepackage{float}
    - \newcommand{\beginsupplement}{ 
        \setcounter{table}{0}  
        \renewcommand{\thetable}{S\arabic{table}} 
        \setcounter{figure}{0} 
        \renewcommand{\thefigure}{S\arabic{figure}}
      }

@@

Corresponding author:

@@

@@

All network analyses were performed using the igraph package [v. 1.1.2; https://igraph.org/r/; @G.2006] in R [v. 3.4.1; https://www.r-project.org/; @RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify significant connections. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were set to 0 (all other edges retained their respective z-values). The proportion of remaining edges (i.e. network density) was high for all individuals (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`), and the retained edges mostly consisted of positive correlations (mean proportion positive = `r mPos`, sd = `r sdPos`). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution within the set of retained edges. Non-retained edges were reset to 0 to exclude them from the network analyses. We generated and analyzed network weight matrices at four levels: (1) for each subject's full concatenated dataset (up to 4800 TRs); (2) on each step of a sliding window analysis (see Partition Evaluation for more details); (3) for the concatenated time series for the two runs on each day (up to 2400 TRs); and (4) for each run separately (up to 1200 TRs).

### *Community Detection*

@@

A considerable amount of meta-analytic work has been dedicated to pinpointing the spatial embedding of behavioral constructs in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across psychological phenomena [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, the effects elicited by the subjective value of an option and the default network (DN) have been suggested to be inseparable in this area, an overlap that partially extends to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. This apparent neural overlap has important implications, as it has motivated theoretical proposals about ways in which these seemingly dissimilar cognitive domains might recruit a shared set of core cognitive processes [@Acikalin2017; @Northoff2012].

However, the interpretation of overlap in group-level data depends on the degree to which functional organization is heterogeneous across individuals. Recent studies have shown that heteromodal brain regions have considerable variability in functional connectivity across individuals [@Mueller2013], individual-specific functional topography can be occluded in aggregative estimations [@Braga2017; @Gordon2017; @Michalka2015; @Tobyne2018], and overlap in functional activation can vanish with increases in spatial precision [@Woo2014]. These findings suggest that group-level and meta-analysis-level overlap does not necessarily imply overlap in individual brains. However, our understanding of the individual-level heterogeneity in the functional topography of mPFC has been mostly descriptive so far. Therefore, affirming that among DN and valuation share neural substrates in this area first requires: 1) a method to reliably and precisely capture its functional topography in isolated individuals; and 2) quantifying the degree of heterogeneity of its intrinsic functional organization in a large group of individuals.

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For 100 individuals, we reliably subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, including identifying communities deterministically, constraining comunities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values to determine the strength of a node's community affiliation, and diagnosing the connectedness of a network through examination of its resulting eigenvalues [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities [@Garcia2017], suggested the binary partitioning was appropriate.

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. However, the precise spatial organization of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individuals. This idiosyncrasy in functional topography was particularly pronounced in mPFC, and was present with various amounts of data. This degree of individual-specifitcity can arise from a variety of sources. For example, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to systematically vary across individuals [@Mackey2014]. Individual variability could also be due to shifts in functional organization that are independent of structural features [@Haxby2013], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. A good goal for future work would be to assess whether the network layout in this region can be predicted on the basis of aspects of brain structure, such as sulcal morphology, myeloarchitecture [@Glasser2016], or structural connectivity [@Saygin2016].

Network-partitioning methods such as SP are data-driven, and therefore provide no labeling information about the resulting communities. We circumvented this issue by identifying the DN community based on its coverage of area 7m, a region in PCC that was preferentially affiliated to the DN literature in the meta-analysis. We were able to apply labels derived from this group-level approach on the basis of the topography in PCC, where functional organization was more consistent across individuals. Because each community spanned both mPFC and PCC, the labels extended to mPFC where topography was more heterogeneous.

>>>>

---
title: 'Individual heterogeneity in the functional network topography of medial prefrontal cortex'
author: "Claudio Toro-Serey^1^, Sean M. Tobyne^2^, & Joseph T. McGuire^1^"
output:
  pdf_document: 
    keep_tex: true
    highlight: haddock
bibliography: mPFC_paper.bib
csl: apa.csl
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
    - \linenumbers
    - \usepackage{float}
    - \newcommand{\beginsupplement}{ 
        \setcounter{table}{0}  
        \renewcommand{\thetable}{S\arabic{table}} 
        \setcounter{figure}{0} 
        \renewcommand{\thefigure}{S\arabic{figure}}
      }

@@

Corresponding author: Claudio Toro-Serey (ctoro@bu.edu)

@@

All code used in this study is openly available at [https://github.com/ctoroserey/mPFC_partitioning](https://github.com/ctoroserey/mPFC_partitioning)

@@

All network analyses were performed using the igraph package [v. 1.1.2; https://igraph.org/r/; @G.2006] in R [v. 3.4.1; https://www.r-project.org/; @RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify significant connections. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were set to 0 (all other edges retained their respective z-values). The proportion of remaining edges (i.e. network density) was high for all individuals (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`), and the retained edges mostly consisted of positive correlations (mean proportion positive = `r mPos`, sd = `r sdPos`). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution within the set of retained edges. Non-retained edges were reset to 0 to exclude them from the network analyses. We generated and analyzed network weight matrices at four levels: (1) for each subject's full concatenated dataset (up to 4800 TRs); (2) on each step of a sliding window analysis (see Partition Evaluation for more details); (3) for the concatenated time series for the two runs on each day (up to 2400 TRs); and (4) for each run separately (up to 1200 TRs).

### *Community Detection*

@@

A considerable amount of meta-analytic work has been dedicated to characterizing the brain activity patterns associated with psychological processes in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across domains [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, activation patterns associated with subjective valuation and with the default network (DN) have been suggested to be inseparable in this area, with overlap also partially extending to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. This apparent neural overlap has important implications, as it has motivated theoretical proposals about ways in which these superficially dissimilar cognitive domains might recruit a shared set of core cognitive processes [@Acikalin2017; @Northoff2012].

However, the interpretation of overlap in group-level data depends on the degree to which functional organization is heterogeneous across individuals. Recent studies have shown that heteromodal brain regions have considerable variability in functional connectivity across individuals [@Mueller2013], individual-specific functional topography can be occluded in aggregative estimations [@Braga2017; @Gordon2017; @Michalka2015; @Tobyne2018], and overlap in functional activation can vanish with increases in spatial precision [@Woo2014]. These findings suggest that group-level and meta-analysis-level overlap does not necessarily imply overlap in individual brains. However, our understanding of the individual-level heterogeneity in the functional topography of mPFC has been mostly descriptive so far [@Braga2017; @Braga2019; Gordon2017]. To affirm that DN and valuation share neural substrates in this area requires a method to reliably and precisely capture the functional topography of mPFC in isolated individuals, as well as a quantitative estimate of the degree of topographic heterogeneity across a large group of individuals.

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For 100 individuals, we subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use parcels that are several orders of magnitude larger than vertices as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, including identifying communities deterministically, constraining communities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values to determine the strength of a node's community affiliation, and the ability to diagnose the connectedness of a network through examination of its resulting eigenvalues [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities [@Garcia2017], suggested the binary partitioning was appropriate.

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. The precise spatial positioning of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individual. The idiosyncrasy in functional topography was particularly pronounced in mPFC, and was identified in both run-based and day-based analyses. Individual-specificity could theoretically arise from a variety of sources. For example, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to vary systematically across individuals [@Mackey2014; Lopez2019]. Individual variability could also be due to shifts in functional organization that are independent of structural features [@Haxby2013], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. An important goal for future work will be to assess whether the network layout in this region can be predicted on the basis of aspects of brain structure, such as sulcal morphology, myeloarchitecture [@Glasser2016], or structural connectivity [Osher2016; @Saygin2011; @Saygin2016].

Network-partitioning methods such as SP are data-driven, and therefore provide no labeling information about the resulting communities. We circumvented this issue by independently identifying the DN community based on its coverage of area 7m, a region in PCC that was preferentially associated with the DN relative to subjective valuation in our meta-analysis. We were able to apply labels derived from this group-level approach on the basis of the topography in PCC, where functional organization was more consistent across individuals. Because each community spanned both mPFC and PCC, the labels extended to mPFC where topography was more heterogeneous.
